No proof nets for \capsabbrev{mll} with units\\[5pt]\Large Proof equivalence in \capsabbrev{mll} is \capsabbrev{pspace}-complete

### 2014-01-21 main/LICS.tex
terminology: free (edge-gadget) => mobile

\section{Rewiring proof nets}

\aftersection
%
In this section we will explore the global rewiring behaviour of proof nets.
%
We will look at notions of subnets, and we will give a simple account of equivalence for the fragment of \capsabbrev{mll} that omits the par.
%
Furthermore, we will  \emph{parity},

@@

% TODO: mention the parity problem

@@

For this argument we shall assume that the connectives $\tn$ and $\parr$ are n-ary, and that every argument of a $\tn$ is a $\parr$ and vice versa. Assume also that $\Gamma$ consists of a single formula, if necessary by introducing a $\parr$ at the root. Now fix an arbitrary switching of $\Gamma$; this switching together with a linking $\links$ defines a tree whose vertices are the subformulae of $\Gamma$ and whose edges are the edges of the chosen switching together with the links of $\links$. Take the root of the formula $\Gamma$ to be the root of this tree. Now consider the edges of the tree to be directed away from the root; so the root has in-degree $0$ and the other vertices have in-degree $1$. This therefore establishes a bijection between the edges of the tree and the non-root vertices.
%
A second linking $\links'$ similarly establishes such a bijection. Given both linkings $\links$ and $\links'$, we compose these bijections to obtain a permutation of the proper subformulae of $\Gamma$. Finally the parity of the pair of linkings is taken to be the parity of this permutation.

@@

These characteristics hold the following lessons on rewiring for the design of our encoding, in Section~\ref{sec:encoding}. 
%
Firstly, making proof nets equivalent is generally easier than preventing them from being equivalent, and in particular changing a formula $A$ to $A\tn\bot$ dramatically increases its rewiring potential.
%
The focus of our encoding will therefore be on preventing undesired linkings from being proof nets at all, rather than being unreachable by rewiring.
%
Secondly, the problem that proof nets up to a single exchange of links are generally not equivalent is too subtle to be useful as a feature of an encoding, and shall mean no more than a technical distraction.

@@

\smallskip
A \emph{reconfiguration step} $\gamma\perm*\delta$ relates two (partial) configurations for an \capsabbrev{ncg} $G$ that differ in value (or definedness) on exactly one edge.
%
The reflexive--transitive closure of $(\perm*)$ will be denoted $(\perm)$.
%
\end{definition}

@@

\aftersection
%
In this section we will establish that in our encoding of constraint graphs %partial \capsabbrev{ncg}-reconfiguration 
in \capsabbrev{mll}, reconfiguration on graphs is captured by proof equivalence in \capsabbrev{mll}.
%
Up to minor technical details, it will be shown that $\itn\gamma\perm\itn\delta$ if and only if $\gamma\perm\delta$, in two parts: completeness ($\Leftarrow$) and soundness ($\Rightarrow$).

% TODO: mention the parity problem

@@

In a proof net for $\itn G$, an edge-gadget $\itn e$ is \emph{well linked} if: 1) its indicator connects to the indicator target of an appropriate vertex-gadget $\itn v$, and 2) each weight element is either in a standard linking with a constraint element of $\itn v$, or linked only to weight absorbers.
%
A well-linked edge-gadget is \emph{free} if all its weight elements connect to weight absorbers.
%
In the following main lemma we will exchange the weight element linking to a constraint element.
%
The lemma considers a sequent consisting of just a vertex-gadget, some appropriate edge-gadgets, and sufficiently many weight absorbers.

@@

\begin{lemma}
\label{lem:permute weight absorbers}
If $\links$ is a well linked proof net for $\itn G$ with a free edge-gadget $\itn e$, and $\links'$ agrees with $\links$ up to an even permutation of weight absorbers, then $\links\perm\links'$.
\end{lemma}

\begin{IEEEproof}
Let $e_0$ be the indicator of $\itn e$; since $\itn e$ is free, $e_0$ may re-attach anywhere within the proof net.
%
Let $e_1$ and $e_2$ be arbitrary other $\bot$-occurrences in $\itn e$.

@@

Let $\gamma$ and $\delta$ agree on every edge except $e$, where $\gamma(e)=v$ and $\delta(e)=w$.
%
Firstly, using Lemma~\ref{lem:octopus roll}, for the edges $d$ other than $e$ such that $\gamma(d)=v$, the weight elements of the edge-gadgets $\itn d$ may be linked to the constraint elements of $\itn v$, in accordance with the target configuration $\itn\delta$.
%
Since $e$ is mobile in $\gamma$, the weights of the edges $d$ suffice to fill the inflow constraint of $v$, and correspondingly the weight elements of edge-gadgets $\itn d$ suffice to fill the constraint elements of $\itn v$, so that $\itn e$ is free.
%
Next, the indicator vertex of $\itn e$, which links to the indicator target of $\itn v$, is re-attached to the indicator target of $\itn w$.
%
Again using Lemma~\ref{lem:octopus roll}, the weight elements of edge-gadgets connected to $\itn w$, including $\itn e$, may be linked in accordance with $\itn\delta$.
%
The resulting proof net is $\itn\delta$ modulo a permutation of weight absorbers; then it is equivalent to either $\itn\delta$ or $\itn\delta'$ by Lemma~\ref{lem:permute weight absorbers}.
\end{IEEEproof}

>>>>

\section{Rewiring proof nets}
\label{sec:rewiring}

\aftersection
%
In this section we will explore the global rewiring behaviour of proof nets.
%
We will look at notions of subnets, and we will give a simple account of equivalence for the fragment of \capsabbrev{mll} that omits the par.
%
Furthermore, we will introduce a notion of relative \emph{parity} between nets, which if odd, guarantees inequivalence.

@@

@@

For this argument we shall assume that the connectives $\tn$ and $\parr$ are n-ary, and that every argument of a $\tn$ is a $\parr$ and vice versa. Assume also that $\Gamma$ consists of a single formula, if necessary by introducing a $\parr$ at the root. Now fix an arbitrary switching of $\Gamma$; this switching together with a linking $\links$ defines a tree whose vertices are the subformulae of $\Gamma$ and whose edges are the edges of the chosen switching together with the links of $\links$. Take the root of the formula $\Gamma$ to be the root of this tree.
%
Now consider the bijection between non-root vertices and edges that associates each vertex with the edge towards its parent. 
%
A second linking $\links'$ similarly establishes such a bijection. Given both linkings $\links$ and $\links'$, we compose these bijections to obtain a permutation of the non-root vertices in $\Gamma$. Finally the parity of the pair of linkings is taken to be the parity of this permutation.
%
% Now consider the edges of the tree to be directed away from the root; so the root has in-degree $0$ and the other vertices have in-degree $1$. This therefore establishes a bijection between the edges of the tree and the non-root vertices.
%
%A second linking $\links'$ similarly establishes such a bijection. Given both linkings $\links$ and $\links'$, we compose these bijections to obtain a permutation of the proper subformulae of $\Gamma$. Finally the parity of the pair of linkings is taken to be the parity of this permutation.

@@

These characteristics hold the following lessons on rewiring for the design of our encoding, in Section~\ref{sec:encoding}. 
%
Firstly, making proof nets equivalent is generally easier than preventing them from being equivalent, and in particular changing a formula $A$ to $A\tn\bot$ dramatically increases its rewiring potential.
%
The focus of our encoding will therefore be on preventing undesired linkings from being proof nets at all, rather than being unreachable by rewiring.
%
Secondly, the problem that proof nets up to a single exchange of links are generally not equivalent is too subtle to be useful as a feature of an encoding.
%
It will prove a minor technical distraction, that we will call the `parity problem'.

@@

\smallskip
A \emph{reconfiguration step} $\gamma\perm*\delta$ relates two (partial) configurations for $G$ that differ in value (or definedness) on exactly one edge; this edge is then called \emph{mobile} in $\gamma$ and $\delta$.
%
The reflexive--transitive closure of $(\perm*)$ will be denoted $(\perm)$.
%
\end{definition}

@@

\aftersection
%
In this section we will establish that in our encoding of constraint graphs %partial \capsabbrev{ncg}-reconfiguration 
in \capsabbrev{mll}, reconfiguration on graphs is captured by proof equivalence in \capsabbrev{mll}.
%
It will be shown that $\gamma\perm\delta$ if and only if $\itn\gamma\perm\itn\delta$ (modulo the parity problem---see Section~\ref{sec:rewiring}).
%
This will be separated into two parts: completeness ($\Rightarrow$) and soundness ($\Leftarrow$) .

@@

In a proof net for $\itn G$, an edge-gadget $\itn e$ is \emph{well linked} if: 1) its indicator connects to the indicator target of an appropriate vertex-gadget $\itn v$, and 2) each weight element is either in a standard linking with a constraint element of $\itn v$, or linked only to weight absorbers.
%
A well-linked edge-gadget is \emph{mobile} if all its weight elements connect to weight absorbers.
%
In the following main lemma we will exchange the weight element linking to a constraint element.
%
The lemma considers a sequent consisting of just a vertex-gadget, some appropriate edge-gadgets, and sufficiently many weight absorbers.

@@

\begin{lemma}
\label{lem:permute weight absorbers}
If $\links$ is a well linked proof net for $\itn G$ with a mobile edge-gadget $\itn e$, and $\links'$ agrees with $\links$ up to an even permutation of weight absorbers, then $\links\perm\links'$.
\end{lemma}

\begin{IEEEproof}
Let $e_0$ be the indicator of $\itn e$; since $\itn e$ is mobile, $e_0$ may re-attach anywhere within the proof net.
%
Let $e_1$ and $e_2$ be arbitrary other $\bot$-occurrences in $\itn e$.

@@

Let $\gamma$ and $\delta$ agree on every edge except $e$, where $\gamma(e)=v$ and $\delta(e)=w$.
%
Firstly, using Lemma~\ref{lem:octopus roll}, for the edges $d$ other than $e$ such that $\gamma(d)=v$, the weight elements of the edge-gadgets $\itn d$ may be linked to the constraint elements of $\itn v$, in accordance with the target configuration $\itn\delta$.
%
Since $e$ is mobile in $\gamma$, the weights of the edges $d$ suffice to fill the inflow constraint of $v$, and correspondingly the weight elements of edge-gadgets $\itn d$ suffice to fill the constraint elements of $\itn v$, so that $\itn e$ is mobile.
%
Next, the indicator vertex of $\itn e$, which links to the indicator target of $\itn v$, is re-attached to the indicator target of $\itn w$.
%
Again using Lemma~\ref{lem:octopus roll}, the weight elements of edge-gadgets connected to $\itn w$, including $\itn e$, may be linked in accordance with $\itn\delta$.
%
The resulting proof net is $\itn\delta$ modulo a permutation of weight absorbers; then it is equivalent to either $\itn\delta$ or $\itn\delta'$ by Lemma~\ref{lem:permute weight absorbers}.
\end{IEEEproof}
